---
title: "tidy-model"
author: "Lam Tran"
date: "2025-02-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library('tidyverse') ; library('modeldata')
```
## Introduction

This R Markdown document demonstrates my abilities to use models for data analysis using a data set collected on crickets.

```{r load data, include = FALSE}
data(crickets, package = "modeldata")
names(crickets)
```

## What is that sound?

In this report, we examine what factors a cricket's chirp rate.

```{r summary, echo = FALSE}
summary(crickets)
ggplot(crickets, aes(x = rate)) +
  geom_histogram(bins = 8) + 
  ggtitle("Distribution of the chirp rate of crickets") +
  xlab('Chirp rate (per min.)')
```

The number of observations in the data set is 31. The number of species is 2. The temperate range is a minimum of 17.20 and a maximum of 30.40. The mean rate of chirping is 72.89.

## Temperature affects chirp rate

```{r temp, echo= FALSE}
ggplot(crickets, aes(x = temp, y = rate)) +
  geom_point() + 
  geom_smooth(method = 'lm') +
  ggtitle("Plot of temperature and chirp rate") +
  ylab('Chirp rate (per min.)') +
  xlab('Temperature (Celsius)')

temp_lm <- lm(rate ~ temp, crickets)

summary.lm(temp_lm)
```

Based on a scatter plot of temperature and chirping and a correlation test, it seems that as temperature increases one degree, the rate of chirping increases about 4.2 chirps per minute.

## Species-specific effects of temperature on chirping

```{r species, echo= FALSE}
ggplot(crickets, aes(x = temp, y = rate, color = species)) +
  geom_point() + 
  geom_smooth(method = 'lm') +
  ggtitle("Plot of temperature and chirp rate for two species of crickets") +
  ylab('Chirp rate (per min.)') +
  xlab('Temperature (Celsius)')
  
  
species_lm <- lm(rate ~ temp + species, crickets)

summary.lm(species_lm)
```

The second graph shows a distinction between the species compared to each other based on the same variables of Chirp Rate and Temperature. To be more explicit, we can see the O. exclamationis have a higher rate of chirp against the other variable of temperature than the O. niveus. 